\ab{Constraints on the quasi-normal mode frequencies of the LIGO-Virgo signals by making full use of gravitational-wave modeling}

### 2021-01-11 papers/intro_paper/intro_paper.tex
changes in Sec. I and II

\documentclass[twocolumn,prd,superscriptaddress,amsfonts,amssymb,amsmath,preprintnumbers]{revtex4-1}

\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{times}
\usepackage{xspace}
\usepackage[varg]{txfonts}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
%\linenumbers

@@

\newcommand{\h}{\mathpzc{h}}
\newcommand{\Hhat}{\hat{\mathpzc{H}}}
\newcommand{\B}{\mathpzc{B}}
\newcommand{\hlm}{\mathpzc{h}_{\ell m}}
\newcommand{\xilm}{\xi_{\ell m}}
\newcommand{\Ylm}{{Y}^{-2}_{\ell m}}
\newcommand{\Y}{{Y}^{-2}}
\newcommand{\hc}{h_\times}
\newcommand{\hp}{h_+}
\newcommand{\Fc}{F_\times}
\newcommand{\Fp}{F_+}
\newcommand{\Mf}{M_f}
\newcommand{\cA}{\mathpzc{A}}
\newcommand{\lm}{_{\ell m}}
\newcommand{\deff}{d_\mathrm{eff}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bxi}{\bm{\xi}}
\newcommand{\bxigr}{\bm{\xi}_{\text{GR}}}
\newcommand{\bxingr}{\bm{\xi}_{\text{nGR}}}
\newcommand{\bzeta}{\bm{\zeta}}
\newcommand{\bs}[1]{\bm{\vec{s}_{#1}}}
\newcommand{\Mo}{M_{\odot}}
\newcommand{\FFe}{\mathrm{FF}_\mathrm{eff}}
\newcommand{\FF}{\mathrm{FF}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\rhoopt}{\rho_\mathrm{opt}}
\newcommand{\rhosubopt}{\rho_\mathrm{subopt}}
\newcommand{\fqnm}{f}
\newcommand{\sigmaqnm}{\sigma}
\newcommand{\n}{\mathbf{n}}
\newcommand*{\skymapscale}{0.5}
\newcommand*{\paramestscale}{0.455}
\newcommand{\df}[1]{\delta f_{\text{#1}}}
\newcommand{\dtau}[1]{\delta \tau_{\text{#1}}}
\newcommand{\fngr}[1]{f_{\text{#1}}}
\newcommand{\taungr}[1]{\tau_{\text{#1}}}
\newcommand{\fgr}[1]{f ^{\text{GR}}_{\text{#1}}}
\newcommand{\taugr}[1]{\tau ^{\text{GR}}_{\text{#1}}}
\newcommand{\pSEOB}{\texttt{pSEOBNR}}
\newcommand{\SEOB}{\texttt{SEOBNR}}

@@

\ab{The observation of GWs has} had significant astrophysical and cosmological
implications~\citep{LSC_2016astroph,gw170817_mma,gw170817_joint,gw170817_hubble}. It
has also allowed us to \sout{make statements in} \ab{probe} fundamental
physics \sout{.Specifically, LIGO-Virgo's observations have allowed us to} and 
test predictions of Einstein's theory of General Relativity 
(GR)~\citep{GR} in the previously unexplored \ab{highly-dynamical, and strong field 
regime} \sout{of highly relativistic, strong-field regimes of
gravity}~\citep{LSC_2016grtests,GW170817_TGR,gwtc1_tgr}. In GR\sout{, a CBC
involving two BHs (}a binary black hole (BBH) system\sout{)} is described \sout{in} \ab{by}
three \sout{distinct} phases: an early \textit{inspiral}, where the two
compact objects spiral in \ab{loosing energy because of the emission of GWs} \sout{and \textit{plunge} due to a back-reaction
of gravitational-wave-emission}, a \textit{merger} marked by the
formation of a\sout{n} \ab{common} apparent horizon~\citep{NRpaper}, and a \sout{late-time} 
\textit{ringdown}, \sout{where} \ab{during which} the newly formed remnant object settles down
to a \sout{stable} Kerr \ab{BH} \sout{state through the emission of} \ab{emitting} an exponentially damped
quasi-normal-mode (QNM) spectrum of gravitational
radiation~\citep{vishu,earlyqnmpapers}.

The LIGO-Virgo collaborations \ab{have released companion papers \cite{}} detailing their \sout{latest} results 
of tests of GR using \sout{events} \ab{GW150914 \cite{} and 
the first two transient catalogues (TC) (i.e., GWTC-1\cite{}  and GWTC-2 \cite{})}. 
The results include tests of GW generation and source dynamics, where bounds are placed on
parametrized deviations in the post-Newtonian \ab{(PN)} coefficients describing
the early inspiral~\citep{earlydevelopmentpapers}, and
phenomenological coefficients describing the intermediate (plunge) and
merger regimes of coalescence~\citep{TIGERmethodspapers}; tests of GW
propagation, which assume a generalized dispersion relation and place
upper bounds on the Compton wavelength and consequently, the mass of
the graviton~\citep{gw170104,samajdar2017projected}, and tests of the
polarization of gravitational radiation using a
multi-GW-detector network~\citep{gw170814,isi2017probing}. The \ab{GWTC-1/2} paper\ab{s} also 
check\sout{s} for consistency between different portions of the signal using estimates
\sout{of} \ab{for the} predicted mass and spin of the remnant
object~\citep{Ghosh:2016xx,Ghosh:2017gfp,LSC_2016grtests}, and
consistency of the residuals with interferometric
noise~\citep{Ghonge:2020suv,gwtc1_tgr}. None of these tests report any
departure from the predictions of GR.

\ab{The first paper on tests of GR by the LIGO Collaboration~\cite{} provided us with the 
first measurement of the dominant damped-sinusoidal signal in the ringdown stage of a BBH 
coalescence.} The LIGO-Virgo O3a testing GR paper~\cite{} \ab{has recently reported} 
\sout{also details results on tests} \ab{a comprehensive analysis of the properties of the 
remnant object, including the ringdown stage, for tens of GW events.}   
\sout{BH ringdown and nature of the remnant object, which is an active field
of research at the moment.} The no-hair conjecture in GR~\citep{}
states that an (electrically neutral) astrophysical BH is completely
described by two observables: mass and spin angular momentum. One
consequence of the no-hair conjecture is that the (complex) QNM
frequencies of gravitational radiation emitted by a perturbed isolated
BH is uniquely determined by its mass and spin angular momentum. Hence
a test of the no-hair conjecture would involve checking for
consistency between estimates of mass and spin of the remnant object
across multiple QNM frequencies~\cite{Dreyer:2003bv}. An inconsistency would either
indicate a non-BH nature of the remnant object, or an incompleteness
of GR as the underlying theory of gravity.

The consistency between the post-merger signal and the least damped QNM was first demonstrated in 
Ref.~\citep{LSC_2016grtests} for GW150914, and later extended to include
overtones in Refs.~\citep{Brito:2018rfr,Giesler:2019uxc,Isi:2019aib,Bhagwat:2019dtm,Forteza:2020hbw}. Consistency
of the late-time \sout{signal} \ab{waveform} with a single QNM is a test of the ringdown of
a BBH coalescence, but not necessarily a test of the no-hair
conjecture, which requires the measurement of (at least) two QNMs (BH
spectroscopy), and checking for consistency between them~\cite{Dreyer:2003bv,Berti:2005ys}. Recent work
in that direction \comment{AB: which direction? There are papers by Nikhef group and 
also Brito et al., and many others that have investigated BH spectroscopy.} include~\citep{Carullo:2018gah,Carullo:2019flw,Bhagwat:2019bwv}. The
nature of the remnant object has also been explored through tests of
BH thermodynamics, like the Hawking's area
theorem~\citep{Cabero:2017avf} or through search for echos in the
post-merger
signal~\citep{Nielsen:2018lkf,Tsang:2019zra,Lo:2018sep,Abedi:2018npz,Abedi:2020sgg,Testa:2018bzd}. None
of these tests have found evidence for non-BH nature of the remnant
object (as described in GR) in LIGO-Virgo BBH observations.

\comment{AB: I don't understand the next senetnce, it looks to me out of context.} 
Most of the tests mentioned above focus on analysing the post-merger
or late-time ringdown signal in isolation. \sout{Second generation
ground-based interferometric detectors like} \ab{Ground-based detectors, 
such as \sout{Advanced} LIGO and Virgo,} are most sensitive to 
stellar-mass BH binaries that merge near the
minim\ab{um}\sout{a} of their sensitivity band ($\sim 100$Hz). As a consequence, the
remnant object \textit{rings down} \comment{AB: we are not discussing the interferometer noises, 
so why to give such a detail about the shot noise?} \sout{in shot-noise dominated higher
frequencies, leaving} \ab{at frequencies where the sensitivity is worse, leaving} 
very little signal-to-noise ratio (SNR) in the
post-merger signal. \comment{AB: This discussion about the starting time of the ringdown starts too abruptly, 
and looks out of context and incomplete. Also, why are we citing only those papers? The matter 
has a long history. There have been calibrations to NR of damped-sinusoid signals.} 
The ringdown start time is not a clearly defined
quantity and has been explored in detail
in Ref.~\citep{Bhagwat:2017tkm}. In other
works~\citep{Carullo:2018gah,Carullo:2019flw}, it has been left as a
free parameter to be estimated directly from the data. Uncertainties
in estimates of the ringdown start-time, as well as an overall lack of
SNR in the post-merger signal, given typical sensitivities of
ground-based detectors result in significant statistical uncertainties
in the measurement of the QNM frequencies.

An independent approach to BH spectroscopy, based on the full-signal
analysis, was \sout{outlined} \ab{introduced} in Ref.~\citep{Brito:2018rfr} 
(henceforth referred to as \paperone). Unlike methods that focus \ab{only} 
on the post-merger signal, it \sout{describes a framework in which a} 
\ab{employs the} complete inspiral-merger-ringdown
(IMR) waveform \sout{model is used} to measure the complex QNM
frequencies. This \sout{allows} \ab{gives} access to the full SNR of the signal,
reducing measurement uncertainties. Moreover, the definition of the
ringdown start time is built into the \ab{merger-ringdown} model and does not need to be
either left as an additional free parameter or fixed using alternate
definitions. While \paperone presented the method \ab{and tested it for }\sout{in the context of a
non-spinning waveform model} \ab{non-spinning BBHs}, \ab{here} we extend the analysis to the \ab{more 
realistic astrophysical} case in which BHs \ab{carry} spin\sout{s in the current paper}. \ab{Furthermore, 
the IMR waveform model used in this paper is more accurate than what employed in \paperone, 
because it contains more information from PN theory and it was calibrated to a much larger set 
of numerical-relativity waveforms~\cite{Bohe:2016gbl}.}  All astrophysical BHs are expected to be spinning, 
and ignoring effects of spin have been shown to introduce systematic biases in the measurement of the 
source properties \cite{paper_showing_systematics_from_ignoring_spin}.

The rest of the paper is organized as follows. Section~\ref{sec:model} describes our parametrized IMR
waveform model. In Sec.~\ref{sec:method}, we define our testing GR  framework in which we use that model 
to measure the complex QNM frequencies within a Bayesian formalism. Then, in Sec.~\ref{sec:results}, 
we demonstrate our method on real GW events, as well as simulated signals. \ab{In particular, we analyze 
the ... and constrain the QNM frequecies ...} Finally, we provide a summary of
our results and discuss future developments in Sec.~\ref{sec:discussion}.

@@

A GW signal from the (quasi-circular) coalescence of two BHs is completely described in GR by a fixed set of $15$ parameters, $\bxigr$. These can be grouped into the \emph{intrinsic} parameters: the masses, $m_1, m_2$ and spins, $\bs1, \bs2$ of the initial binary and a reference time and phase $t_c, \phi_c$ respectively; and the \emph{extrinsic} parameters: the sky position of the binary ($\alpha$, $\delta$), the luminosity distance, $d_L$, and its orientation described through the inclination of the binary $\iota$ and its polarisation $\psi$. If the spins of the individual BHs are assumed to be aligned with the orbital angular momentum of the binary, then GW signal has a reduced dimensionality of $11$ parameters. The aligned-spin components are defined as: $a_{iz} = s_{iz}/m_i$, where $i=1,2$ for the two BHs.

@@

As in \paperone, we use an IMR waveform model developed within the EOB formalism. However, in this work we extend the results of \paperone by using a waveform model for spinning, non-precessing BBHs. Namely we use as our baseline model the aligned-spin multipolar waveform developed in Ref.~\citep{Cotesta:2018fcv}. In addition to being calibrated to NR simulations, this model also uses information from BH perturbation theory for the merger and ringdown phases. Therefore this waveform model provides a very accurate and faithful semi-analytic description of the inspiral, merger and ringdown phases. Henceforth we will denote this model by $\SEOB$ for short.

@@

where $\varphi$ is the azimuthal direction to the observer, ${}_{-\!2}Y_{\ell m}(\theta,\phi)$ are the $-2$ spin-weighted spherical harmonics. The $\SEOB$ model we employ includes the $(\ell, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$ modes~\cite{Cotesta:2018fcv}. For each $(\ell, m)$, the inspiral-(plunge-)merger-ringdown $\SEOB$ waveform is schematically given by

@@

where $\theta(t)$ is the Heaviside step function, $h_{\ell m}^\mathrm{insp-plunge}$ represents the inspiral-plunge part of the waveform, whereas $h_{\ell m}^\mathrm{merger-RD}$ denotes the merger-ringdown phase.   
The merger-ringdown $\SEOB$ modes read~\citep{Bohe:2016gbl,Cotesta:2018fcv}

@@

are fixed by imposing that the functions $\tilde{A}_{\ell m}(t)$ and $\tilde{\phi}_{\ell m}(t)$ are of class $C^1$ at $t = t_{\textrm{match}}^{\ell m}$, when matching the merger-ringdown waveform to the inspiral-plunge $\SEOB$ waveform $h_{\ell m}^\mathrm{inspiral-plunge}(t)$. This allows to write the coefficients $c_{i,c}^{\ell m}$ as~\cite{Cotesta:2018fcv}:

@@

@@

@@

@@

In the $\SEOB$ model constructed in Ref.~\cite{Cotesta:2018fcv}, the complex frequencies $\sigma_{\ell m 0}$  are expressed in terms of the final BH mass and spin~\cite{Berti:2005ys,Berti:2009kk}, and the latter are related to the BBH's component masses and spins through NR--fitting-formulas obtained in GR~\cite{Taracchini:2013rva,Hofmann:2016yih}. Here instead, in the spirit of what was done in \paperone, we promote the QNM (complex) frequencies to be free parameters of the model, while keeping the inspiral-plunge modes $h_{\ell m}^\mathrm{inspiral-plunge}(t)$ fixed to their GR values. Henceforth called $\pSEOB$, this model uses a parameterised version of the $\SEOB$ model where the frequency and the damping time of the ${\ell m 0}$ mode, i.e, $(f_{\ell m 0}, \tau _{\ell m 0})$ is defined through fractional deviations, ($(\delta f_{\ell m 0},\delta \tau_{\ell m 0})$), from the corresponding GR values, i.e., the predictions of the QNM frequencies obtained using the fits~\cite{Taracchini:2013rva,Hofmann:2016yih}:

\begin{eqnarray}
f_{\ell m 0} &:=& f_{\ell m 0}^{\text{GR}} (1 + \delta f_{\ell m 0})\label{eq:nongr_freqs_a} \\ 
\tau _{\ell m 0} &:=& \tau _{\ell m 0}^{\text{GR}} (1 + \delta \tau_{\ell m 0}) \label{eq:nongr_freqs_b}
\end{eqnarray}
The GR quantities $( f_{\ell m 0}^{GR},\tau_{\ell m 0}^{GR})$ are constructed using the same NR--fitting--formula prescriptiion described above. We note that when leaving $\sigma_{\ell m}$ to vary freely, the functions $\tilde{A}_{\ell m}(t)$ and $\tilde{\phi}_{\ell m}(t)$ will in general also differ from the GR prediction, since those functions depend on the QNM complex frequencies, as can be seen from the expressions for $c_{i,c}^{\ell m}$ and $d_{1,c}^{\ell m}$.

\subsection{Bayesian PE}\label{sec:method}

The parametrised model, $\pSEOB$, described above introduces an additional set of non-GR parameters, $\bxingr = (\delta f_{\ell m 0},\delta \tau_{\ell m 0})$, corresponding to each $(\ell,m)$ QNM present in the GR waveform model $\SEOB$. One then proceeds to use the Bayes theorem to obtain the \emph{posterior} probability distribution on $\blambda = \{\bxigr, \bxingr\}$, given a hypothesis $\mathcal{H}$:

>>>>

%\documentclass[twocolumn,prd,superscriptaddress,amsfonts,amssymb,amsmath,preprintnumbers,nofootinbib]{revtex4-1}
\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,eqsecnum,amsfonts,amsmath]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{times}
\usepackage{xspace}
\usepackage[varg]{txfonts}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
%\linenumbers

@@

\newcommand{\h}{\mathpzc{h}}
\newcommand{\Hhat}{\hat{\mathpzc{H}}}
\newcommand{\B}{\mathpzc{B}}
\newcommand{\hlm}{\mathpzc{h}_{\ell m}}
\newcommand{\xilm}{\xi_{\ell m}}
\newcommand{\Ylm}{{Y}^{-2}_{\ell m}}
\newcommand{\Y}{{Y}^{-2}}
\newcommand{\hc}{h_\times}
\newcommand{\hp}{h_+}
\newcommand{\Fc}{F_\times}
\newcommand{\Fp}{F_+}
\newcommand{\Mf}{M_f}
\newcommand{\cA}{\mathpzc{A}}
\newcommand{\lm}{_{\ell m}}
\newcommand{\deff}{d_\mathrm{eff}}
\newcommand{\rmi}{\mathrm{i}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bxi}{\bm{\xi}}
\newcommand{\bxigr}{\bm{\xi}_{\text{GR}}}
\newcommand{\bxingr}{\bm{\xi}_{\text{nGR}}}
\newcommand{\bzeta}{\bm{\zeta}}
\newcommand{\bs}[1]{\bm{\vec{S}_{#1}}}
\newcommand{\Mo}{M_{\odot}}
\newcommand{\FFe}{\mathrm{FF}_\mathrm{eff}}
\newcommand{\FF}{\mathrm{FF}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\rhoopt}{\rho_\mathrm{opt}}
\newcommand{\rhosubopt}{\rho_\mathrm{subopt}}
\newcommand{\fqnm}{f}
\newcommand{\sigmaqnm}{\sigma}
\newcommand{\n}{\mathbf{n}}
\newcommand*{\skymapscale}{0.5}
\newcommand*{\paramestscale}{0.455}
\newcommand{\df}[1]{\delta f_{\text{#1}}}
\newcommand{\dtau}[1]{\delta \tau_{\text{#1}}}
\newcommand{\fngr}[1]{f_{\text{#1}}}
\newcommand{\taungr}[1]{\tau_{\text{#1}}}
\newcommand{\fgr}[1]{f ^{\text{GR}}_{\text{#1}}}
\newcommand{\taugr}[1]{\tau ^{\text{GR}}_{\text{#1}}}
\newcommand{\pSEOB}{\texttt{pSEOBNR}}
\newcommand{\SEOB}{\texttt{SEOBNR}}

@@

\ab{The observation of GWs has} had significant astrophysical and cosmological
implications~\citep{LSC_2016astroph,gw170817_mma,gw170817_joint,gw170817_hubble}. It
has also allowed us to \sout{make statements in} \ab{probe} fundamental
physics \sout{.Specifically, LIGO-Virgo's observations have allowed us to} and 
test predictions of Einstein's theory of General Relativity 
(GR)~\citep{GR} in the previously unexplored \ab{highly-dynamical, and strong field 
regime} \sout{of highly relativistic, strong-field regimes of
gravity}~\citep{LSC_2016grtests,GW170817_TGR,gwtc1_tgr}. In GR\sout{, a CBC
involving two BHs (}a binary black hole (BBH) system\sout{)} is described \sout{in} \ab{by}
three \sout{distinct} phases: an early \textit{inspiral}, where the two
compact objects spiral in \ab{loosing energy because of the emission of GWs} \sout{and \textit{plunge} due to a back-reaction
of GW-emission}, a \textit{merger} marked by the
formation of a\sout{n} \ab{common} apparent horizon~\citep{NRpaper}, and a \sout{late-time} 
\textit{ringdown}, \sout{where} \ab{during which} the newly formed remnant object settles down
to a \sout{stable} Kerr \ab{BH} \sout{state through the emission of} \ab{emitting} \sout{an exponentially damped} 
quasi-normal-modes (QNMs) \sout{spectrum of gravitational
radiation}~\citep{vishu,earlyqnmpapers} \ab{(i.e., damped oscillations with specific, discrete 
frequencies and decay times)}.

The LIGO-Virgo collaborations \ab{have released companion papers \cite{}} detailing their \sout{latest} results 
of tests of GR  \sout{using events} \ab{for GW150914} \cite{} and \ab{for several GW events of} 
the \sout{first} two transient catalogues (TC): GWTC-1\cite{}  and GWTC-2 \cite{}. 
The results include tests of GW generation and source dynamics, where bounds are placed on
parameterized deviations in the post-Newtonian \ab{(PN)} coefficients describing
the early inspiral~\citep{earlydevelopmentpapers}, and
phenomenological coefficients describing the intermediate (plunge) and
merger regimes of coalescence~\citep{TIGERmethodspapers}; tests of GW
propagation, which assume a generalized dispersion relation and place
upper bounds on the Compton wavelength and, consequently, the mass of
the graviton~\citep{gw170104,samajdar2017projected}, and tests of the
polarization of gravitational radiation using a
multi--GW-detector network~\citep{gw170814,isi2017probing}. The \ab{GWTC-1/2} paper\ab{s} also 
check\sout{s} for consistency between different portions of the signal using estimates
\sout{of} \ab{for the} predicted mass and spin of the remnant
object~\citep{Ghosh:2016xx,Ghosh:2017gfp,LSC_2016grtests}, and
consistency of the residuals with interferometric
noise~\citep{Ghonge:2020suv,gwtc1_tgr}. None of these tests report any
departure from the predictions of GR.

\ab{The first paper on tests of GR by the LIGO Collaboration~\cite{} provided us with the 
first measurement of the dominant damped-oscillation signal in the ringdown stage of a BBH 
coalescence.} The LIGO-Virgo O3a testing GR paper~\cite{} \ab{has recently reported} 
\sout{also details results on tests} \ab{a comprehensive analysis of the properties of the 
remnant object, including the ringdown stage, for tens of GW events.}   
\sout{BH ringdown and nature of the remnant object, which is an active field
of research at the moment.} The no-hair conjecture in GR~\citep{}
states that an (electrically neutral) astrophysical BH is completely
described by two observables: mass and spin angular momentum. One
consequence of the no-hair conjecture is that the (complex) QNM
frequencies of gravitational radiation emitted by a perturbed isolated
BH \ab{are} \sout{is} uniquely determined by its mass and spin angular momentum. Hence
a test of the no-hair conjecture would involve checking for
consistency between estimates of mass and spin of the remnant object
across multiple QNM frequencies~\cite{Dreyer:2003bv}. An inconsistency would either
indicate a non-BH nature of the remnant object, or an incompleteness
of GR as the underlying theory of gravity.

The consistency between the post-merger signal and the least damped QNM was first demonstrated in 
Ref.~\citep{LSC_2016grtests} for GW150914, and later extended to include
overtones in Refs.~\citep{Brito:2018rfr,Giesler:2019uxc,Isi:2019aib,Bhagwat:2019dtm,Forteza:2020hbw}. Consistency
of the late-time \sout{signal} \ab{waveform} with a single QNM is a test of the ringdown of
a BBH coalescence, but not necessarily a test of the no-hair
conjecture, which requires the measurement of (at least) two QNMs (BH
spectroscopy), and \sout{checking for} consistency between them~\cite{Dreyer:2003bv,Berti:2005ys}. Recent work
in that direction \comment{AB: which direction? There are papers by Nikhef group and 
also Brito et al., and many others that have investigated BH spectroscopy.} include~\citep{Carullo:2018gah,Carullo:2019flw,Bhagwat:2019bwv}. The
nature of the remnant object has also been explored through tests of
BH thermodynamics, like the Hawking's area
theorem~\citep{Cabero:2017avf} or through search for echos in the
post-merger
signal~\citep{Nielsen:2018lkf,Tsang:2019zra,Lo:2018sep,Abedi:2018npz,Abedi:2020sgg,Testa:2018bzd}. None
of these tests have found evidence for non-BH nature of the remnant
object (as described in GR) in LIGO-Virgo BBH observations.

\comment{AB: I don't understand the next sentence, it looks to me out of context.} 
Most of the tests mentioned above focus on analysing the post-merger
or late-time ringdown signal in isolation. \sout{Second generation
ground-based interferometric detectors like} \ab{Ground-based detectors, 
such as \sout{Advanced} LIGO and Virgo,} are most sensitive to 
stellar-mass BH binaries that merge near the
minim\ab{um}\sout{a} of their sensitivity band ($\sim 100$Hz). As a consequence, the
remnant object \textit{rings down} \comment{AB: we are not discussing the interferometer noises, 
so why to give such a detail about the shot noise?} \sout{in shot-noise dominated higher
frequencies, leaving} \ab{at frequencies where the sensitivity is worse, leaving} 
very little signal-to-noise ratio (SNR) in the
post-merger signal. \comment{AB: This discussion about the starting time of the ringdown starts too abruptly, 
and looks out of context and incomplete. Also, why are we citing only those papers? The matter 
has a long history. There have been calibrations to NR of damped- oscillation signals.} 
The ringdown start time is not a clearly defined
quantity and has been explored in detail
in Ref.~\citep{Bhagwat:2017tkm}. In other
works~\citep{Carullo:2018gah,Carullo:2019flw}, it has been left as a
free parameter to be estimated directly from the data. Uncertainties
in estimates of the ringdown start-time, as well as, an overall lack of
SNR in the post-merger signal, given typical sensitivities of
ground-based detectors, result in significant statistical uncertainties
in the measurement of the QNM frequencies.

An independent approach to BH spectroscopy, based on the full-signal
analysis, was \sout{outlined} \ab{introduced} in Ref.~\citep{Brito:2018rfr} 
(henceforth referred to as \paperone). Unlike methods that focus \ab{only} 
on the post-merger signal, it \sout{describes a framework in which a} 
\ab{employs the} complete inspiral-merger-ringdown
(IMR) waveform \sout{model is used} to measure the complex QNM
frequencies. This \sout{allows} \ab{gives} access to the full SNR of the signal,
reducing measurement uncertainties. Moreover, the definition of the
ringdown start time is built into the \ab{merger-ringdown} model and does not need to be
either left as an additional free parameter or fixed using alternate
definitions. While \paperone presented the method \ab{and tested it for }\sout{in the context of a
non-spinning waveform model} \ab{non-spinning BBHs}, \ab{here} we extend the analysis to the \ab{more 
realistic astrophysical} case in which BHs \ab{carry} spin\sout{s in the current paper}. \ab{Furthermore, 
the IMR waveform model used in this paper is more accurate than what employed in \paperone, 
because it contains higher-order corrections in PN theory and it was calibrated to a much larger set 
of numerical-relativity (NR) waveforms~\cite{Bohe:2016gbl}.}  All astrophysical BHs are expected to be spinning, 
and ignoring effects of spin has been shown to introduce systematic biases in the measurement of the 
source properties \cite{paper_showing_systematics_from_ignoring_spin}.

The rest of the paper is organized as follows. Section~\ref{sec:model} describes our parameterized IMR
waveform model. In Sec.~\ref{sec:method}, we define our framework to test GR\ab{, notably how we can measure 
complex QNM frequencies with our parameterized model within a Bayesian formalism}. Then, in Sec.~\ref{sec:results}, 
we demonstrate our method on real GW events, as well as simulated signals. \ab{In particular, we analyze 
the ... and constrain the QNM frequecies ... Finally, in Sec.~\ref{sec:discussion} we provide a summary of
our results and discuss future developments.}

@@

\comment{AB: In teh following, I will not use much the strike-out, but instead I will 
bracket in color the text that I have changed.}

A GW signal from the (quasi-circular) coalescence of two BHs is
completely described in GR by a \sout{fixed set of} $15$ parameters,
$\bxigr$. These can be grouped into the \emph{intrinsic} parameters:
the masses, $m_1, m_2$ and spins, $\bs1, \bs2$ of the \ab{component 
objects in the} \sout{initial} binary; and the 
\emph{extrinsic} parameters: \ab{a reference time $t_c$ and phase 
$\phi_c$,} the sky position of the binary ($\alpha$,
$\delta$), the luminosity distance, $d_L$, and \ab{the binary's} orientation
described through the inclination of the binary $\iota$ and its
polarization $\psi$. \ab{We also introduce the total mass $M = m_1+m_2$, 
and the (dimensionless) symmetric mass ratio $\nu = m_1m_2/M^2$}.

\ab{Here, we focus on BHs with spins aligned or anti-aligned 
with the orbital angular momentum (henceforth, ``aligned-spin''). In this case,  
the GW signal depends on $11$ parameters. We denote the 
aligned-spin \ab{(dimensionless)} components as $\chi_{i} = |\vec{\bm{S}}_i|/m^2_i$, where $i=1,2$ for the two BHs.}

@@

As in \paperone, we use an IMR waveform model developed within the effective-one-body (EOB) 
formalism~\cite{}. \ab{However, whereas \paperone was limited to non-spinning multipolar waveforms, 
here we use as our baseline model the aligned-spin multipolar waveform model 
developed in Ref.~\citep{Cotesta:2018fcv}}. In addition to being
calibrated to NR simulations, this model also uses information from BH
perturbation theory for the merger and ringdown phases. \sout{Therefore this
waveform model provides a very accurate and faithful semi-analytic
description of the inspiral, merger and ringdown phases.} \ab{Henceforth we
will denote this model by $\SEOB$ for short}~\footnote{\ab{In the LIGO Algorithm Library (LAL), this 
waveform model is called {\tt SEOBNRv4HM}.}}.

@@

\comment{AB: there is a mismatch of notations for the parameters with respect to what was introduced above.}
where $\varphi$ is the azimuthal direction to the observer, ${}_{-\!2}Y_{\ell m}(\theta,\varphi)$ are the $-2$ spin-weighted spherical harmonics. The $\SEOB$ model we employ includes the $(\ell, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$ modes~\cite{Cotesta:2018fcv}. For each $(\ell, m)$, the inspiral-(plunge-)merger-ringdown $\SEOB$ waveform is schematically given by

@@

where $\theta(t)$ is the Heaviside step function, $h_{\ell m}^\mathrm{insp-plunge}$ represents the inspiral-plunge part of the waveform, whereas $h_{\ell m}^\mathrm{merger-RD}$ denotes the merger-ringdown \ab{waveform, which reads~\citep{Bohe:2016gbl,Cotesta:2018fcv}}

@@

are fixed by imposing that the functions $\tilde{A}_{\ell m}(t)$ and $\tilde{\phi}_{\ell m}(t)$ are of class $C^1$ at $t = t_{\textrm{match}}^{\ell m}$, when matching the merger-ringdown waveform to the inspiral-plunge $\SEOB$ waveform $h_{\ell m}^\mathrm{inspiral-plunge}(t)$. This allows us to write the coefficients $c_{i,c}^{\ell m}$ as~\cite{Cotesta:2018fcv}:

@@

\label{c1}

@@

\label{c2}

@@

\label{d1}    

@@

In the $\SEOB$ model constructed in Ref.~\cite{Cotesta:2018fcv}, the
complex frequencies $\sigma_{\ell m 0}$ are expressed in terms of the
final BH mass and spin~\cite{Berti:2005ys,Berti:2009kk}, and the
latter are related to the BBH's component masses and spins through
NR--fitting-formulas obtained in
GR~\cite{Taracchini:2013rva,Hofmann:2016yih}. Here instead, in the
spirit of what was done in \paperone, we promote the QNM (complex)
frequencies to be free parameters of the model, while keeping the
inspiral-plunge modes $h_{\ell m}^\mathrm{inspiral-plunge}(t)$ fixed
to their GR values. \ab{More explicitly, we introduce a parameterized 
version of the $\SEOB$ model where the frequency and the
damping time of the ${\ell m 0}$ mode (i.e, $(f_{\ell m 0}, \tau
_{\ell m 0})$) is defined through the fractional deviations, $(\delta
f_{\ell m 0},\delta \tau_{\ell m 0})$, from the corresponding GR
values, which are obtained using NR fits~\cite{Taracchini:2013rva,Hofmann:2016yih}. Thus,} 
\begin{eqnarray}
f_{\ell m 0} &=& f_{\ell m 0}^{\text{GR}}\, (1 + \delta f_{\ell m 0})\,,\label{eq:nongr_freqs_a} \\ 
\tau _{\ell m 0} &=& \tau _{\ell m 0}^{\text{GR}}\, (1 + \delta \tau_{\ell m 0})\,. \label{eq:nongr_freqs_b}
\end{eqnarray}
\ab{We denote such a parameterized waveform model \pSEOB}~\footnote{\ab{This 
waveform model is called {\tt pSEOBNRv4HM} in LAL.}}.

\ab{As said}, the GR quantities $( f_{\ell m 0}^{GR},\tau_{\ell m 0}^{GR})$ are
constructed using the same NR--fitting--formula \sout{prescriptiion} 
described above. We note that when leaving $\sigma_{\ell m}$ to vary
freely, the functions $\tilde{A}_{\ell m}(t)$ and $\tilde{\phi}_{\ell
  m}(t)$ \sout{will} in general also differ from the GR prediction\ab{s}, since
those functions depend on the QNM complex frequencies, as can be seen
from the expressions for $c_{i,c}^{\ell m}$ and $d_{1,c}^{\ell m}$ \ab{in Eqs.~(\ref{c1}), 
(\ref{c2}), and (\ref{d1}).}

\subsection{Bayesian Parameter Estimation}\label{sec:method}
